---
title: "Regression Analysis in R"
author: "David Griswold"
date: "8/6/2018"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
library(tidyverse)
library(caret)
```


# Linear

## What is Linear Regression?

Regression is using a line and least squares methodology to estimate future values off of a current dataset.

## When to use it

**We should use linear regression to predict for a continous Y varible**

In order to build a valid model, we need to control for the folowing asumptions inhernt in the model:

1. Linear relationship
2. Error terms that are uncorrelated and have non-constant variance
3. Independent predictor varibles

## The Model 

Poupulation Version:
$$ Y = \beta_0 + \beta_{1}X_{1} + e $$

Sample Version:
$$ y = b_0 + b_{1}x_{1} + e_{i} $$

## Code It

Let's predict the price of a diamond.

```{r}
ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
  geom_point()
```

## Model It!
```{r}
diamonds <- mutate(diamonds, Price_Sqrt = sqrt(price))

diamondModel <- lm(price ~ carat, data = diamonds)
summary(diamondModel)

qplot(x = diamonds$carat, y = diamondModel$residuals)
```

## Dummy Varibles

So what if we want to use a set of varibles that are not a continous numeric set? We can use a dummy varilbe! If its coded as a factor, R basically handels it for us

```{r}
diamondModel <- lm(Price_Sqrt ~ carat + cut, data = diamonds)
summary(diamondModel)


View(model.matrix(diamondModel))
```


# Logistic Regression 

Sometimes we want to use regression to take a look at something that isn't a linear predictor, it's a one or a zero...like our titanic dataset. To do that, we have to use a sigmoid function to make them linear. Welcome to logistic regression.

## Read In Data
```{r}
titanic <- read_csv("../titanic.csv")

titanic %<>% mutate(Sex = as.factor(Sex), Embarked = as.factor(Embarked), Survived = as.factor(Survived))
```

## Partition Data
```{r}
index <- sample(1:nrow(titanic), nrow(titanic)*.8)

train <- titanic[index,]

test <- titanic[-index,]
```

## Fit Model
```{r}
titanicModel <- glm(Survived ~ Sex + Age, family = binomial("logit"), data = train)



summary(titanicModel)
```

## Let's Make a Perdiction! 
```{r}
prediction <- predict(titanicModel, newdata = select(test,-Survived)) 

prediction <-  ifelse(prediction > 0.5,1,0) %>% as.factor()

confusionMatrix(prediction, test$Survived)
```

