---
title: "Dplyr + GGPlot"
author: "David Griswold"
date: "7/20/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Review: Group By and Summarise

## Why is this important

- A lot of time we have individual data and questison about the whole.
- Examples:
  - User sessions data but quesitons about the site's expierence
  - Data by the week and questions about the year
  - Data on the passenger but questions about the titanic
- Using this functionality let's us bridge those gaps efficently

## Pratice It
### Read Data in
```{r}
library(tidyverse)
library(magrittr)

titanicData <- read.csv("../titanic.csv")
```


## Breif Digression Using NyC Flights Dataset

Sometimes, we are dealing with a response that is continous rather than binary like the titanic dataset. R has a few that are baked in.

```{r}
flightData <- nycflights13::flights

# Factorize where necessary
flightData$carrier <- as.factor(flightData$carrier)
flightData$origin <- as.factor(flightData$origin)
```


## On your own:

- What is the average departure delay?
- Which airport is the busiest? (Group by and summarise)
- Which day is busiest?


### Why do factor's matter?

It let's the software know if the data is catagorical or not. This is really important for a lot of the graphical plots as well as doing summarization.


## Who has the most flights out of NYC?
```{r}
group_by(flightData, carrier) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

## Which Airports are the most busy?
```{r}
group_by(flightData, origin) %>% 
  summarise(count = n())
```

## When are most fligths taking off?
```{r}
library(lubridate)
flightData$time_hour <- ymd_hms(flightData$time_hour) 

group_by(flightData, hour = hour(time_hour)) %>% 
  summarise(flights = n()) %>% 
  ggplot() + # we can tack GGPLOT right onto the end with another pipe operator
  geom_line(aes(x = hour, y = flights))
```

## Which time of day is busiest each week?
```{r}
group_by(flightData, weekday = wday(time_hour), hour = hour(time_hour)) %>% 
  summarise(flights = n()) %>% 
  ggplot() +
  geom_line(aes(x = hour, y = flights)) +
  facet_wrap(~ weekday)


```


## How about on the same graph?
```{r}

```


# ggplot: Back to Titanic

## Raw Survival Rates
```{r}
titanicData$Survived %<>% as.factor()

ggplot(titanicData) +
  geom_bar(aes(x = Survived, y = ..count../sum(..count..), fill = Survived)) +
  labs(x = "Survived", y = "Proportion", title = "Titanic Survival Rate")
```

## How did survival rates change with Pclass?

What do you think? Save this as `SurvivalxClass`
```{r}
# Really Annoying, but here is how we can go back to numeric 
titanicData$Survived %<>% as.character() %>% as.numeric()

titanicData$Pclass %<>% as.factor()

SurvivalxClass <- group_by(titanicData, Pclass) %>% 
  summarize(Survived = mean(Survived))
```

Let's visualize it! 
```{r}
ggplot(SurvivalxClass) +
  geom_col(aes(x = Pclass, y = Survived, fill = Pclass))
```

If the varible we are working off is a factor, it will fill with that. Two main Bar Chart geom:

- `geom_col` gives us the data as it is in the dataset
- `geom_bar` gives you the density by the factor

## Facets: How did our male/female survival change by pclass?

```{r}
titanicData$Survived %<>% as.character() %>% as.integer()

group_by(titanicData, Pclass, Sex) %>% 
  summarise(count = mean(Survived)) %>% 
  ggplot() +
  geom_col(aes(x = Sex, y = count, fill = Sex)) +
  facet_grid(. ~ Pclass)
```

## Correlation Plots

What is a correlation?

- Way of gauging how much variation is explained by some data
$$\frac{Var_x \times Var_y}{n-1} $$
$$ \frac{Cov_{xy}}{(Var_x)(Var_y)}$$

What percent of the individual variance is explained by the convariance? It ranges between negative one and positive one.

## How do we calculate it?

```{r}
# First, we need to subset to only numeric varibles
subset <- select(titanicData, Survived, SibSp, Parch, Fare)

# Make Correlation
correlation <- cor(subset)

# Plot
library(corrplot)
corrplot(correlation)
```

Not super insightful here. 
