---
title: "Lecture Five"
author: "David Griswold"
date: "7/26/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
titanicData <- read.csv("titanic.csv")
```

# ggplot: Back to Titanic

## Raw Survival Rates
```{r}
titanicData$Survived %<>% as.factor()

ggplot(titanicData) +
  geom_bar(aes(x = Survived, y = ..count../sum(..count..), fill = Survived)) +
  labs(x = "Survived", y = "Proportion", title = "Titanic Survival Rate")
```

## How did survival rates change with Pclass?

What do you think? Save this as `SurvivalxClass`
```{r}
# Really Annoying, but here is how we can go back to numeric 
titanicData$Survived %<>% as.character() %>% as.numeric()

titanicData$Pclass %<>% as.factor()

SurvivalxClass <- group_by(titanicData, Pclass) %>% 
  summarize(Survived = mean(Survived))
```

Let's visualize it! 
```{r}
ggplot(SurvivalxClass) +
  geom_col(aes(x = Pclass, y = Survived, fill = Pclass))
```

If the varible we are working off is a factor, it will fill with that. Two main Bar Chart geom:

- `geom_col` gives us the data as it is in the dataset
- `geom_bar` gives you the density by the factor

## Facets: How did our male/female survival change by pclass?

```{r}
titanicData$Survived %<>% as.character() %>% as.integer()

group_by(titanicData, Pclass, Sex) %>% 
  summarise(count = mean(Survived)) %>% 
  ggplot() +
  geom_col(aes(x = Sex, y = count, fill = Sex)) +
  facet_grid(. ~ Pclass)
```

## Missingness
```{r}
library(Amelia)
```



## Correlation Plots

What is a correlation?

- Way of gauging how much variation is explained by some data
$$\frac{Var_x \times Var_y}{n-1} $$
$$ \frac{Cov_{xy}}{(Var_x)(Var_y)}$$

What percent of the individual variance is explained by the convariance? It ranges between negative one and positive one.

## How do we calculate it?

```{r}
# First, we need to subset to only numeric varibles
subset <- select(titanicData, Survived, SibSp, Parch, Fare)

# Make Correlation
correlation <- cor(subset)

# Plot
library(corrplot)
corrplot(correlation)
```

Not super insightful here. 

