---
title: 'Lecture Three: Let’s Make Some Graphs'
author: "David Griswold"
date: "7/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/davidgriswold/Desktop/LearnR/LectureThree")
```


# Load Titanic Set
```{r}
titanic <- read.csv("../titanic.csv")
```

# GGPlot: What is it and what will it do for us?
GGPLOT is an professional level piece of graphics software that is universally used to produce professional-level scientific graphics. Used by Fivethirtyeight and others to produce graphics. It takes some practice, but it makes incredible graphics.

**Incredable** documentation can be found here:
- [GGplot Core Documentation](https://ggplot2.tidyverse.org/)
- [R For Data Science Chapter](http://r4ds.had.co.nz/data-visualisation.html#the-layered-grammar-of-graphics)

Also, there is always Google.


## Grammar of Graphics
- System for separating data out into layers

----
# It's Time to Start Thinking Like Analysts
> Central Question  **How Can We Predict Survival on the Titanic?**


## High Level Thinking
Graphical work is often used for exploratory analysis and for getting a feel of the dataset. Let's give that a shot.

We have (I hope) all seen Kate and Leo in their academy award winning performance. What factors do you think might influence your chance of making it off the boat? 

*Fun fact: Titanic was the only movie to gross more after opening when Valentine's day pushed sales. 2nd Highest grossing movie of all times behind Avatar (3rd adjusted for inflation)
*

# Class
```{r}
ggplot(titanic, aes(x = Pclass)) +
  geom_bar() 



```



# How about relationships in the data? Let's look at a correlation plot
## Corrplot Package
