---
title: "Basic R Programming"
author: "David Griswold"
date: "June 19, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Why Learn R?

- Scaleable way to clean, manipulate, and model
- Access to ecosystem of reporting tools

## How to get the most out of these lunch and learns
- Code along
- Be super curious
- Ask me anything

## What is a Data Type?
- **When you get an error in your code, it will _very_ often be due to a data type issues**
- A set of rules for storing data with a specific set of tradeoffs 

## The Vector and the Varible
```{r echo=TRUE}
ourVar <- "Hello World"
anotherVar <- 5.5
someOtherVar <- 6
aLogicalVar <- TRUE 
yetAnotherVar <- c("Detroit", "is", "for", "unicorns", "!") 
```

- In R, `<-` is the same thing as `=` when we define varibles
- R sees everything as a vector regardless of lenght

### What is a vector?
- One dimensional array of data
- Can be between length one or very long
- In R, a single piece of information is a vector of length one

## Vector Data Types
```{r, echo = TRUE}
typeof(ourVar)
typeof(anotherVar)
typeof(someOtherVar)
typeof(aLogicalVar)
typeof(yetAnotherVar)
```
- *Character:* Text data `"Hello World"`
- *Integer:* Whole numbers
- *Double:* Floating point numbers
- *Logical:* TRUE and False




