---
title: "Basic R Programming"
author: "David Griswold"
date: "June 19, 2018"
output: 
  ioslides_presentation:
    logo: ../globalStatic/logo.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What is and why use R?
> - Statistical computing language used by companies like [Facebook, Microsoft, NYT, Uber, and many more](https://www.listendata.com/2016/12/companies-using-r.html)
> - Open source community (CRAN)
> - Great visualization and reporting (Like these slides!)
> - R Studio environment


## Why Learn R?
> - Scaleable way to clean, manipulate, visualize, and model data
> - Valuable, marketable skill
> - Opportunity to automate routine data tasks
> - Access to ecosystem of machine learning and AI tools
> - Easier to learn for data than Python (Point of much debate)

## Best Practices for Learning a New Programming Language?
> - Do it, make mistakes
> - Learn to love debugging (as much as one can:)
> - Be curious with documentation
> - Ask me anything and set up extra time to talk
> - Be persistent
> - Give me lots of session feedback

## R Studio?
- Super popular R-focused (but not exclusive) IDE
- Pulls a set of tools together to help us be more productive
- Serves as a hub for the R community
-  **Do we all have it downloaded??**

# Basic R Programming|Data Types and the Dataframe


## Goals for Today
We will build an understanding of:

1. What R is (Hopefully done)
2. How R stores data
3. How to load a .csv into R
4. How to access dataframe information

## What is a Varible?
- A variable is a way to store information for our later use
- In R, we use `<-` to define variables instead of `=` 
  - Technically we can use both, but it is recommended to stick with `<-`
```{r echo=TRUE}
someVar <- "Hello!"
someVar
```

## What is a Data Type?
- A Data Type is a particular set of rules for storing data with a specific set of trade offs and permissible operations for each
- **When you get an error in your code, it will _very_ often be due to a data type issues**

### Base Data Types
- *Character:* Text data, often called a string   
- *Integer:* Whole numbers   
- *Double:* Floating point numbers 
- *Logical:* TRUE and FALSE      
- *Factor:* More complex type for storing categorical data that is a blend of an integer and a string

**All other data types are built on top of these**

## Base Data Types
```{r echo=TRUE}
ourVar <- "Hello World"; typeof(ourVar)
anotherVar <- 5.5; typeof(anotherVar)
someOtherVar <- 6L; typeof(someOtherVar)
aLogicalVar <- TRUE; typeof(aLogicalVar)
```

## Vector
- We need to be able to store more than one value, so R gives us the vector, a one dimensional array of information. 
- The combined command, `c()`, command helps us build a vector of any of the datatypes above
```{r echo=TRUE}
ourVector <- c(3.3, 4.4,6.7,1.1)
length(ourVector)
typeof(ourVector)
```

## Vector
- In fact, R sees even single length pieces of data as a vector of length one
```{r echo=TRUE}
is.atomic(ourVar)
length(ourVar)
```
- **Vectors must all be the same type**
- This gives us speed advantages at scale

## The Matrix
- What if we need to store more than one dimension of data?
- Luckily, we have the matrix
- All data in a matrix must be the same type
- Basically, think of it as a two dimensional vector
```{r echo=TRUE}
ourMatrix <- matrix(c(1.4, 2.2, 3.6, 4.9), c(2,2))
ourMatrix; typeof(ourMatrix)
```

## The Dataframe
- But most of our data has more than one type?
- Enter: the dataframe
- Basically a collection of vectors in a single object

## Rmarkdown
- Make a project folder
- Let's Get started

## Breif Aside on Operators
- There are a wide array of basic operators that R can use like `+`, `-`, `/`, `//`, `%%`
- Also, R has a wide array of logical operators like `>`, `<`, `==`, `%in%`, and others
- We also have the `NA`, or the type for missing data
- We will run into all of these, so we will spend more time on it then