---
title: "Basic R Programming"
author: "David Griswold"
date: "June 19, 2018"
output: 
  ioslides_presentation:
    logo: ../globalStatic/logo.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What is and why use R?
> - Statistical computing languge used by comapnies like [Facebook, Microsoft, NYT, Uber, and many more](https://www.listendata.com/2016/12/companies-using-r.html)
> - Open source community (CRAN)
> - Great visualization and reporting (Like these slides!)
> - R Studio enviorment


## Why Learn R?
> - Scaleable way to clean, manipulate, visualize, and model data
> - Valuable, marketable skill
> - Opporunity to automate routine data tasks
> - Access to ecosystem of machine learning and AI tools
> - Easier to learn for data than Python (Point of much debate)

## Best Practices for Learning a New Programming Language
> - Do it, make mistakes
> - Be curious with documentation
> - Ask me anything and set up extra time to talk
> - Be paitent and be persistant
> - Give me lots of session feedback

## R Studio
- Super poupular R-focused IDE
- Pulls a set of tools together to help us be more productive
- Serves as a hub for the R community
- *Do we all have it downloaded??* 

## What we are going to learn
1. Basics of programming and data reading (Base R)
2. Titanic data exploration using (Dplyr, ggplot)
3. Data cleaning and manipulation (stringr, dplyr)
4. Correlation and additional visualization (ggplot, corrplot)
5. Linear modeling (lm(), MASS)
6. K-Means cluster (caret)

This is all very subject to change. This will be my first time setting up a course, so please let me know how to improve.

# Basic R Programming

# Goal for Today
We have an understanding of:
1. What R is
2. How R stores information
3. How to load a .csv into R
4. How to access dataframe information

## How does computer memory work?
*Draw pointer diagram*

## What is a Data Type?
- Why is this important? **When you get an error in your code, it will _very_ often be due to a data type issues**
- What is it? 
  - A set of rules for storing data with a specific set of tradeoffs to each

## The Vector and the Varible
```{r echo=TRUE}
ourVar <- "Hello World"
anotherVar <- 5.5
someOtherVar <- 6
aLogicalVar <- TRUE 
yetAnotherVar <- c("Detroit", "is", "for", "unicorns", "!") 
```

- In R, `<-` is the same thing as `=` when we define varibles
- R sees everything as a vector regardless of lenght

### What is a vector?
- One dimensional array of data
- Can be between length one or very long
- In R, a single piece of information is a vector of length one

## Vector Data Types
- *Character:* Text data `"Hello World"`
- *Integer:* Whole numbers `3`
- *Double:* Floating point numbers `3.3`
- *Logical:* TRUE and False `TRUE`
- *Factor:* 

Why are vectors important?
> - Allows us to store and manipulate a lot of data very efficently giving a speed advantage at scale.

## Vector Data Types
```{r, echo = TRUE}
typeof(ourVar)
typeof(anotherVar)
typeof(someOtherVar)
typeof(aLogicalVar)
typeof(yetAnotherVar)
```

## Breif Aside on Operators
> - There are a wide array of basic operators that R can use like `+`, `-`, `/`, `//`, `%%`
> - Also, R has a wide array of logical operators like `>`, `<`, `==`, `%in%`, and others
> - We also have the `NA`, or the type for missing data.
    - Any time you use `NA` anything that it touches will go to `NA`

# Rmarkdown
- Make a project folder
- Let's Get started

# The Data Frame and Reading .csv Data

## Matrix No More
- Matrix came before dataframe. Allowed us to store two-dimensional data, but only of the same type.
- Data frames alow us to store square data of differnt types 
- Think of a Dataframe as a collection of lists at a particular chunk of memory

## Let's Load a Dataframe
```{r echo=TRUE}
titanicData <- read.csv("../titanic.csv")
```





