---
title: 'Lecture 1: Load Titanic Dataset'
author: "David Griswold"
date: "June 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Load Titanic Set
```{r}
titanic <- read.csv("titanic.csv")
```

# Head of Titanic Set
```{r}
# Some Title
head(titanic)
```
The `head()` functions allow us to view the first five rows of a dataframe.

# Structure and Types of the Titanic Set
```{r}
str(titanic)
```
The `str()` gives us the data set's structure, or the variables and their respective types. This is important because they are often wrong and require changing.

# Dimensions of the Dataset
```{r}
dim(titanic)
```
The `dim()` function gives us the dimensions as a vector. (Always rows then columns). We can also access these independently with `nrow()` and `ncol()`.

# Indexing
```{r}
mean(titanic$Survived)

titanic[5,]
```
To index, we can use a few different commands. Each routinely comes in handy. 
- To grab a vector, use `$`. *This returns a vector*
- To grab a specific row, we can use the rc notation and []. *This returns a view of the dataframe*
- We can do the same to get a column
- Vectors are indexed the same way!!

# Change Data Type
```{r}
titanic$Name <- as.character(titanic$Name)
```
Using what we learned so far, let's select a piece of memory and update the data type. 
- Any time we want to change the type, we will use a as.desiredtype() command and an assignment operator.

# Update a Piece of Information
```{r}

```
Use what we know so far to update a cell. Let's change a name... Remember Mr. Allen. Let's make sure Jack Dawson is well accounted for. 

# Drop a Column
```{r}
titanic$Ticket <- NULL
```
Use `NULL` and what we have learned to drop a column.
- If we ever want to remove something from global memory, we can use `rm()`
- If we ever want to clear it, click the broom. 

# Add a New Column
```{r}
titanic$Fam <- titanic$Parch + titanic$SibSp
```
- Use what we know to write in a new one
- If we want to overwrite a new one 

# Overwrite a Column
```{r}

```
Use assignment to overwrite a column with another piece of data

